---
title: "Data Preprocessing"
author: "Freddy A. Julkanain II"
output: pdf_document
---
Tidy will be mainly use for wrangling, bestNormalize for effective data transformation, and MASS package for some basic statistical graphs, test, and manipulations.
```{r}
pacman::p_load(
  tidyverse,
  bestNormalize,
  MASS
  )
```

Loading the given raw data into the environment.
```{r}
DF=read_csv("radiomics_completedata.csv")
```

Simple check for any missing value
```{r}
sum(is.na(DF))
```

Isolating numeric columns, removing the target classification variable and applying kolmogorov-smirnov test for normality to each column since shapiro test performs badly for n<50.
```{r,warning=F}
DF1=DF%>%select_if(is.numeric)
DF1=DF1[,-1]
DF2=apply(DF1,2,function(x){ks.test(x,"pnorm")})
```
Exttacting pvalue from test results via lapply and converting result to an atomic vector for easier manipuation
```{r}
KS_list=unlist(lapply(DF2, function(x) x$p.value))
```

Using the p-value vector, well create a simple table indicating number of columns that would be considered nonnormal using 0.05 as region of rejection.
```{r}
nonnormal=sum(KS_list<0.05)
Normal=sum(KS_list>=0.05)
ColNumber=ncol(DF1)
cbind(ColNumber,nonnormal,Normal)
```


We will now apply the orderNorm transformation fuction from bestNormalize fuction which does an Ordinary Quantile Normalization which is a very strong normalization technique. The result of the transformation is extracted namely x.t and then collected to a dataframe right before again applying the normality test algorithm previously 
```{r,warning=F}
tmpDF=DF[,c(3:length(names(DF)))]%>%
  apply(2,orderNorm)%>%
  lapply( function(x) x$x.t)%>%
  as.data.frame()
KS_list=apply(tmpDF,2,function(x){ks.test(x,"pnorm")})
KS_list=unlist(lapply(KS_list, function(x) x$p.value))
```
Redoing the table for column normality check we see that all nonnormal columns are transformed into nonnormal data.
```{r,warning=F}
nonnormal=sum(KS_list<0.05)
Normal=sum(KS_list>=0.05)
ColNumber=ncol(DF1)
cbind(ColNumber,nonnormal,Normal)
```
The transformation results are loaded back into the main dataframe in the environment
```{r}
DF[,c(3:length(names(DF)))]=tmpDF
```

A basic correlation heatmap is generated( excluding the first two columns that are categorical). It is technically too compressed due to the size of the data
```{r}
CorMatrix=cor(DF[,-c(1,2)])
heatmap(CorMatrix,Rowv=NA,Colv=NA,scale="none",revC = T)
```

Lastly, the data is written into a CSV file for easier access of following model buildings.
```{r}
write.csv(DF,"CleanedDF.csv")
```




