---
title: "MODEL - 3"
author: "Freddy A. Julkanain II"
output: pdf_document 
---

# Helper Packages And Modeling Packages
```{r}
library(dplyr)    
library(ggplot2)   
library(stringr)  
library(cluster)    
library(factoextra)
library(gridExtra)  
library(tidyverse)
library(readr)
library(mclust)
```

## Load the dataset
#Note that the data we used here is from the output of Reprocessing Rawdata
```{r}
DF = read_csv("CleanedDF.csv")

```

# Scaling/Standardizing the Data
```{r}
df = scale(DF[c(3:431)])

```
#QUESTION: Compare the following clustering technique results:
# 1. K-means
# 2. Hierarchical 
# 3. Model Based


#------------------------------------ K-MEANS CLUSTERING ------------------------------------------#
```{r}
kmeans(df, centers = 3, iter.max = 100, nstart = 100)
clusters = kmeans(df, centers = 3, iter.max = 100, nstart = 100)
```


# determine and visualize optimal number of clusters
```{r}
fviz_nbclust(df, kmeans, method = "wss") 
fviz_nbclust(df, kmeans, method = "silhouette")
fviz_nbclust(df, kmeans, method = "gap_stat") 

clusters = kmeans(df, centers = 2, iter.max = 100, nstart = 100)
fviz_cluster(kmeans(df, centers = 2, iter.max = 100, nstart = 100), data = df)
```


#The quality of a k-means partition. The quality of the partition is
```{r}
clusters$betweenss / clusters$totss
```


# visualize clusters using original variables
```{r}
clusters = kmeans(df, centers = 3, iter.max = 100, nstart = 100)
DF = DF |> mutate(cluster = clusters$cluster)
DF |> ggplot(aes(x = Failure, y = Entropy_cooc.W.ADC, col = as.factor(cluster))) + geom_point()
```

# 2.Heirarchical Clustering
```{r}
FPD = DF%>%
  select_if(is.numeric) %>%  # select numeric columns
  select(-Failure.binary) %>%    # remove target column
  mutate_all(as.double) %>%  # coerce to double type
  scale()
data = dist(FPD, method = "euclidean")

```

# Hierarchical clustering using Complete Linkage
```{r}
hc1 = hclust(data, method = "complete")
plot(hc1, cex = 0.6)
rect.hclust(hc1, k = 2, border = 1:4)
```
#AGNES
```{r}
set.seed(123)
hc2 = agnes(FPD, method = "complete")
hc2$ac
```
#DIANA
```{r}
hc4 = diana(FPD)
hc4$dc

p1 = fviz_nbclust(FPD, FUN = hcut, method = "wss", 
                   k.max = 10) +
  ggtitle("(A) Elbow method")
p2 = fviz_nbclust(FPD, FUN = hcut, method = "silhouette", 
                   k.max = 10) +
  ggtitle("(B) Silhouette method")
p3 = fviz_nbclust(FPD, FUN = hcut, method = "gap_stat", 
                   k.max = 10) +
  ggtitle("(C) Gap statistic")
gridExtra::grid.arrange(p1, p2, p3, nrow = 1)
```
# Ward's method
```{r}
hc5 = hclust(data, method = "ward.D2" )
sub_grp = cutree(hc5, k = 8)
table(sub_grp)
```




# 3.Model Based
```{r}
mod3 = Mclust(df[,1:10], G=3) 
summary(mod3)
mod4 = Mclust(df, 1:9) #from sir lecture 

#Error in plot.new() : figure margins too large
summary(mod4)

```
# Plot results
```{r}
plot(mod3, what = "density") #cannot plot 1:428 #dugay pag 1:10
plot(mod3, what = "uncertainty")
```


```{r}
legend_args = list(x = "bottomright", ncol = 5)
plot(mod3, what = 'BIC', legendArgs = legend_args)
plot(mod3, what = 'classification')
plot(mod3, what = 'uncertainty')
```


```{r}
probabilities = mod3$z 
colnames(probabilities) = paste0('C', 1:3)

probabilities = probabilities %>%
  as.data.frame() %>%
  mutate(id = row_number()) %>%
  tidyr::gather(cluster, probability, -id)

ggplot(probabilities, aes(probability)) +
  geom_histogram() +
  facet_wrap(~ cluster, nrow = 2)
```


```{r}
uncertainty = data.frame(
  id = 1:nrow(df),
  cluster = mod3$classification,
  uncertainty = mod3$uncertainty
)
uncertainty %>%
  group_by(cluster) %>%
  filter(uncertainty > 0.25) %>%
  ggplot(aes(uncertainty, reorder(id, uncertainty))) +
  geom_point() +
  facet_wrap(~ cluster, scales = 'free_y', nrow = 1)
```



```{r}
cluster2 = df %>%
  scale() %>%
  as.data.frame() %>%
  mutate(cluster = mod3$classification) %>%
  filter(cluster == 2) %>%
  select(-cluster)

cluster2 %>%
  tidyr::gather(product, std_count) %>%
  group_by(product) %>%
  summarize(avg = mean(std_count)) %>%
  ggplot(aes(avg, reorder(product, avg))) +
  geom_point() +
  labs(x = "Average standardized consumption", y = NULL)
```



